Implementation and comparison of register allocation to translate \lstinline!LLVM--! to x86

### 2023-11-27 article.tex
More detail

\documentclass{article}
\usepackage[style=ieee]{biblatex}

@@

  \draw [decorate,decoration={brace,amplitude=5pt,mirror,raise=4ex}]
    (0.3,-1) -- (0.3,1) node[midway,xshift=3em,rotate=-90]{Backend};
  \draw [decorate,decoration={brace,amplitude=5pt,mirror,raise=4ex}]
    (0.3,1.25) -- (0.3,3.25) node[midway,xshift=3em,rotate=-90]{Frontend};
\end{tikzpicture}}
\caption{Compiler phases, backend highlighted}
\end{figure}
Most operations of a general-purpose programming language are translated to a set of logic,  control, and arithmetic processor instructions, that are intended to be executed sequentially on a computer processor: a single circuit/chip, referred to as the \textit{central processing unit} (CPU). CPU design varies and has evolved over time, but most processors executing logic operations are \textit{register machines}, in that they use a limited set of \textit{general-purpose registers} (GPRs) to store working values in combination with \textit{random access memory} (RAM) for mid-term and other I/O peripherals for long-term storage.

The reason for this can largely be attributed to performance, as register machines routinely outperform so-called stack machines \cite{ShiYunhe2008VmsS} that are often used in VMs.
Due to this, and the limited amount of GPRs available, a crucial part of the backend stage for an optimizing compiler is assigning each variable  of the source program to a GPR in such a way that maximizes performance without sacrificing correctness.

The process of assigning each variable to a GPR is generally referred to as \textit{register allocation}, and

>>>>

\documentclass{article}
\usepackage[style=apa]{biblatex}

@@

  \draw [decorate,decoration={brace,amplitude=5pt,mirror,raise=4ex}]
    (0.3,-1) -- (0.3,1) node[midway,xshift=3em,rotate=-90]{Backend};
  \draw [decorate,decoration={brace,amplitude=5pt,mirror,raise=4ex}]
    (0.3,1.25) -- (0.3,3.25) node[midway,xshift=3em,rotate=-90]{Frontend};
\end{tikzpicture}}
\caption{Compiler phases, backend highlighted}
\end{figure}
Most operations of a general-purpose programming language are translated to a set of logic,  control, and arithmetic instructions that are intended to be executed sequentially on a computer processor: a single circuit/chip, referred to as the \textit{central processing unit} (CPU). CPU design varies and has evolved over time, but most processors executing logic operations are \textit{register machines}, in that they use a limited set of \textit{general-purpose registers} (GPRs) to store working values in combination with \textit{random access memory} (RAM) for mid-term and other I/O peripherals for long-term storage.

This can largely be attributed to performance, as register machines routinely outperform so-called \textit{stack machines} \parencite{ShiYunhe2008VmsS} that are often used in VMs.
Because of the limited amount of GPRs available simultaneously, a crucial part of the backend stage for an optimizing compiler is assigning each variable  of the source program to a GPR in such a way that maximizes performance without sacrificing correctness.

The process of assigning each variable to a GPR is referred to as \textit{register allocation}, and can be approached in several different ways. This paper will seek to implement graph coloring and linear scan and evaluate them in terms of runtime performance after compilation. %% TODO: compare with time-complexity/compile-time duration?
The primary sources will be \textit{Modern Compiler Implementation in ML} \parencite{AppelAndrewW} and \textit{Compilers: Principles, techniques, and tools} by \parencite{alma99122774374105763}, in addition to scientific publications concerning the linear scan approach.
